FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-02-27 ess_marmo.tex

\begin{keypoints}
\item Inter-operable data formats 
\item Planetary data standards
\item Data processing and visualization
\end{keypoints}

@@

\subsection{Aladin}
Aladin is both a sky atlas and a portal to access data -- mainly images and
catalogues -- available through the VO.
It is able to load FITS files and to display images taking into account various
WCS projections.
Like DS9, Aladin is popular and broadly used in the part of the astronomical
community manipulating images.
So far it is also the reference generator and visualizer of Hierarchical
Progressive Survey (HiPS).
HiPS\footnote{http://www.ivoa.net/documents/HiPS/} is a recommendation endorsed
by the International Virtual Observatory Alliance (IVOA).
It describes how to access to sky survey data stored at various spatial
resolutions, offering a progressive view of possibly very large datasets.
%The purpose is to allows users to explore all-sky surveys without having to
%explicitly perform any query: the most important information is visible at low
%spatial resolutions while finer details appear by zooming in.
%In the framework of the VESPA project, the idea of using Aladin to display
%planetary data, especially planetary HiPS, has emerged.
HiPS are usually generated from a collection of WCS tagged FITS files.
%As an alternative, Aladin can use JPEG files together with additional metadata
%files.
%A metadata file -- one by JPEG file -- basically contains in a text format the
%WCS part of a FITS header.
Aladin\footnote{http://aladin.u-strasbg.fr/} \citep{aladin} and its HiPS generator have
been updated to support planetary conventions.
%Then, planetary missions usually use different projections when providing
%images of planets polar caps and equatorial region.
%Often the former are in the stereographic (STG) projection while the later
%is in the Plate Carr\'ee (CAR) projection. 
%This specificity has been taken into account into the HiPS generator.
In the future, Aladin will be able to recognize the FITS keywords defined
in the present paper (see Tab. \ref{table:proj} and \ref{table:ctype}).
For example, it will be able to take into account a planet's radii to compute
distances on the surface of the body.
Finally, among the WCS projections listed in Tab. \ref{table:proj}, several of them are
barely used in extrasolar astronomy and are not supported in Aladin yet.
An effort has been started to revamp and update the code of projections in
Aladin.
We expect all projections to be available in a forthcoming release.

>>>>

\begin{keypoints}
\item We present an extension of FITS metadata description for planetary surface
investigations. 
\item A FITS description for planetary data aims
to simplify sharing data across the planetary and astronomy domains.
\item FITS open description metadata will promote interoperability from raw data
formatting to final visualization.
\end{keypoints}

@@

\subsection{Aladin}
Aladin is both a sky atlas and a portal to access data -- mainly images and
catalogues -- available through the VO.
It is able to load FITS files and to display images taking into account various
WCS projections.
Like DS9, Aladin is popular and broadly used in the part of the astronomical
community manipulating images.
So far it is also the reference generator and visualizer of Hierarchical
Progressive Survey (HiPS).
HiPS\footnote{http://www.ivoa.net/documents/HiPS/} is a recommendation endorsed
by the International Virtual Observatory Alliance (IVOA).
It describes how to access to sky survey data stored at various spatial
resolutions, offering a progressive view of possibly very large datasets.
%The purpose is to allows users to explore all-sky surveys without having to
%explicitly perform any query: the most important information is visible at low
%spatial resolutions while finer details appear by zooming in.
%In the framework of the VESPA project, the idea of using Aladin to display
%planetary data, especially planetary HiPS, has emerged.
HiPS are usually generated from a collection of WCS tagged FITS files.
%As an alternative, Aladin can use JPEG files together with additional metadata
%files.
%A metadata file -- one by JPEG file -- basically contains in a text format the
%WCS part of a FITS header.
Aladin\footnote{http://aladin.u-strasbg.fr/} \citep{aladin} and its HiPS generator have
been updated to support planetary conventions (see Figure \ref{fig:aladin}).
%Then, planetary missions usually use different projections when providing
%images of planets polar caps and equatorial region.
%Often the former are in the stereographic (STG) projection while the later
%is in the Plate Carr\'ee (CAR) projection. 
%This specificity has been taken into account into the HiPS generator.

\begin{figure}[ht!]
\centerline{\includegraphics[height=20pc]{ess_marmo6}}
\caption{Visualization of several level of detail on Mars with Aladin: from the
global topographic map, as a HiPS layer, to a high resolution CRISM spectral cube
in FITS format. All spectral analysis tools useful for astronomical spectra
are then available for planetary data.}
\label{fig:aladin}
\end{figure}

In the future, Aladin will be able to recognize the FITS keywords defined
in the present paper (see Tab. \ref{table:proj} and \ref{table:ctype}).
For example, it will be able to take into account a planet's radii to compute
distances on the surface of the body.
Finally, among the WCS projections listed in Tab. \ref{table:proj}, several of
them are barely used in extrasolar astronomy and are not supported in Aladin yet.
An effort has been started to revamp and update the code of projections in
Aladin.
We expect all projections to be available in a forthcoming release.
